---
title: "Lab_5"
author: "Lindley Slipetz"
date: "2/25/2021"
output: html_document
---

### Load packages and data

```{r load-packages, message = FALSE, warning = FALSE}
library(tidyverse) 
#install.packages("wesanderson")
library(wesanderson)

```

```{r load-data, message = FALSE}
load("G:\\My Drive\\DataScience\\Lab_04\\data\\dennys.rda")
load("G:\\My Drive\\DataScience\\Lab_04\\data\\laquinta.rda")
```

### Exercise 1

```{r den_alaska}
dn_ak <- dennys %>%
  filter(state == "AK")
nrow(dn_ak)
```
There are 3 dennys in Alaska.

### Exercise 2
```{r lq_alaska}
lq_ak <- laquinta %>%
  filter(state == "AK")
nrow(lq_ak)
```
There are 2 laquintas in alaska.

### Exercise 3

We need to calculate 6 distances.

### Exercise 4 

```{r alaska_join}
dn_lq_ak <- full_join(dn_ak, lq_ak, by = "state")
dn_lq_ak
```

There are 6 observations in the dataframe. The variables are address, city, state, zip, longitude, latitude for both the x and the y datasets.

### Exercise 5

We use mutate() to add a new variable to the dataframe while keeping the existing variables.

```{r haversine}
haversine <- function(long1, lat1, long2, lat2, round = 3) {
  # convert to radians
  long1 = long1 * pi / 180
  lat1  = lat1  * pi / 180
  long2 = long2 * pi / 180
  lat2  = lat2  * pi / 180
  
  R = 6371 # Earth mean radius in km
  
  a = sin((lat2 - lat1)/2)^2 + cos(lat1) * cos(lat2) * sin((long2 - long1)/2)^2
  d = R * 2 * asin(sqrt(a))
  
  return( round(d,round) ) # distance in km
}

```

###Exercise 6

```{r distance}
dn_lq_ak <- dn_lq_ak %>%
  mutate(distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y, round = 3))
```

### Exercise 7
```{r min_dist}
dn_lq_ak_mindist <- dn_lq_ak %>%
  group_by(address.x) %>%
  summarise(closest = min(distance))
```

### Exercise 8

```{r alaska_plot}
dn_lq_ak_mindist %>%
ggplot(mapping = aes(y = closest, x = reorder(address.x, -closest))) +
  geom_bar(stat="identity") +
  labs(title = "Closest dennys and laquintas", y = "Closest", x = "Address")+
  scale_colour_grey(start = 0, end = .7) +
  theme_bw()
```

```{r alaska_summary}
dn_lq_ak_mindist %>%
  summary(closest)
```

The mean closest distance is 4.410. The minimym is 2.035 and the maximum is 5.998. The distribution has two bars that are nearly equal with one that is much smaller. As I have it arranged, it's skewed to the right, but that's an artifact of how I've arranged the data.
