---
title: "Lab_5"
author: "Lindley Slipetz"
date: "2/25/2021"
output: html_document
---

### Load packages and data

```{r load-packages, message = FALSE, warning = FALSE}
library(tidyverse) 
#install.packages("wesanderson")
library(wesanderson)

```

```{r load-data, message = FALSE}
load("G:\\My Drive\\DataScience\\Lab_04\\data\\dennys.rda")
load("G:\\My Drive\\DataScience\\Lab_04\\data\\laquinta.rda")
```

### Exercise 1

```{r den_alaska}
dn_ak <- dennys %>%
  filter(state == "AK")
nrow(dn_ak)
```
There are 3 dennys in Alaska.

### Exercise 2
```{r lq_alaska}
lq_ak <- laquinta %>%
  filter(state == "AK")
nrow(lq_ak)
```
There are 2 laquintas in alaska.

### Exercise 3

We need to calculate 6 distances.

### Exercise 4 

```{r alaska_join}
dn_lq_ak <- full_join(dn_ak, lq_ak, by = "state")
dn_lq_ak
```

There are 6 observations in the dataframe. The variables are address, city, state, zip, longitude, latitude for both the x and the y datasets.

### Exercise 5

We use mutate() to add a new variable to the dataframe while keeping the existing variables.
